Bandwidth-Aware Reconfigurable Memory Design with Hybrid Memory Technologies

### 2011-04-07 trunk/paper.tex

Memory reconfiguration relies on accurate predictions of demand bandwidth of a
workload in response to the dynamically varying application characteristics.
Conventional last value or table based predictors can model neither long range
patterns of an application nor patterns with variable lengths. In this work, we
employ a statistical predictor to support for the BAROM design. The basic idea
of the predictor is borrowed from the n-Gram models, which are typically used by
speech and natural language processing~\cite{smoothing-tech}. The n-gram models
are usually formulated as a probability distribution $p(s)$ over a strings $s$,
and attemp to reflect how frequently the string occurs as a sentense. Prediction
on a reasonable combination of strings is then generated based on the
probability distribution. Our predictor employs the same basic idea, but
simplifies the model for performance considerations. In our model, each demand
bandwidth (DBW) sample abtained in a time interval is analogous to a string in
language, and a given length (the ``order'') of samples is stored in a table as
a pattern. At each time interval, the prediction engine will update the pattern
table with the new DBW sample, and calculate probability of the updated pattern.
The probability of a pattern $s$ of length $l$ can be calculated without loss as
the product of a sequence of conditional probabilities using the following
equation.

\begin{equation}
  p(s)=p(w_1)p(w_2|w_1)...p(w_l|w_1...w_{l-1})=\prod_{i=1}^{l}p(w_i|w_1...w_{i-1})\label{eq:pr1}
\end{equation}
where $w_l$ is the DBW sample obtained in current time interval, and $w_1$
through $w_{l-1}$ are the preceding $l-1$ DBW samples. In n-gram models, we make
the approximation that each conditional probability only depends on the
preceding $n-1$ samples and obtain Equation~\ref{eq:pr2}.

\begin{equation}
  p(s)=\prod_{i=1}^{l}p(w_i|w_{i-n+1}^{i-1})\label{eq:pr2}
\end{equation}
in which $w_{i-n+1}^{i-1}$ represents the sequence of $w_{i-n+1}...w_{i-1}$.

>>>>

Memory reconfiguration relies on accurate predictions of demand bandwidth of a
workload in response to the dynamically varying application characteristics.
Conventional last value or table based predictors can model neither long range
patterns of an application nor patterns with variable lengths. In this work, we
employ a statistical predictor to support for the BAROM design. The basic idea
of the predictor is borrowed from the n-Gram models, which are typically used by
speech and natural language processing~\cite{smoothing-tech}. The n-gram models
are usually formulated as a probability distribution $p(s)$ over a strings $s$,
and attemp to reflect how frequently the string occurs as a sentense. Prediction
on a reasonable combination of strings is then generated based on the
probability distribution. Our predictor employs the same basic idea, but
simplifies the model for performance considerations. In our model, each demand
bandwidth (DBW) sample abtained in a time interval is analogous to a string in
language, and a given length (the ``order'') of samples is stored in a table as
a pattern. At each time interval, the prediction engine will update the pattern
table with the new DBW sample, and calculate probability of the updated pattern.
The probability of a pattern $s$ of length $l$ can be calculated without loss as
the product of a sequence of conditional probabilities using the equation shown
below.

\begin{equation}
  p(s)=p(w_1)p(w_2|w_1)...p(w_l|w_1...w_{l-1})=\prod_{i=1}^{l}p(w_i|w_1...w_{i-1})\label{eq:pr1}
\end{equation}
where $w_l$ is the DBW sample obtained in current time interval, and $w_1$
through $w_{l-1}$ are the preceding $l-1$ DBW samples. In n-gram models, we make
the approximation that each conditional probability only depends on the
preceding $n-1$ samples and obtain the following equation.

\begin{equation}
  p(s)=\prod_{i=1}^{l}p(w_i|w_{i-n+1}^{i-1})\label{eq:pr2}
\end{equation}
in which $w_{i-n+1}^{i-1}$ denotes the sequence of $w_{i-n+1}...w_{i-1}$. In
order to compute the result of Equation~\ref{eq:pr2}, an estimation of of
$p(w_i|w_{i-n+1}^{i-1})$ can be generated using maximum likelihood~(ML)
estimation method. In n-gram models, the largest n in wide use is n=3 which is
called a trigram model. Therefore, we adopt n=3 in our predictor. Each
conditional probability is calculated using ML estimation in
Equation~\ref{eq:pr3}.

\begin{equation}
  \hat{p}(w_i|w_{i-n+1}^{i-1})=\frac{c(w_{i-2}^{i})}{c(w_{i-2}^{i-1})}\label{eq:pr3}
\end{equation}
where $c(w_{a}^{b})$ is the number of times that the sequence $w_a..w_b$ appears
in preceding samples.
